### 2016-06-29 introduction.tex
added franks intro

%The emergence of powerful and inexpensive single-board computers
%running full versions of a standard \ac{OS} from many possible in the last
%years opens exciting possibilities for researchers. This not only
%allows to seamlessly develop implementations that are compatible with
%higher end devices and allows the reuse stable software from the Open
%Source community. It also enables the deployment and testing of large
%scale (hundreds of devices or more) distributed systems at a fraction
%of the cost necessary in previous years. This paper focuses on
%providing the critical steps and mechanisms followed to deploy, setup,
%and maintain a large-scale testbed based on Raspberry Pi devices of
%different models. We also complement this with an in-depth evaluation
%of the Raspberry Pi as a platform to enable a new paradigm in
%networking and storage known as \ac{NC}, which relies on
%processing at intermediate nodes to achieve higher throughput, delay
%and reliability performance than state-of-the-art routing techniques.
%
%
%% INTRODUCTION TO NETWORK CODING
%Introduced by Ahlswede et al.~\cite{ahlswede2000network}, network
%coding constitutes a paradigm shift in the way how researchers and
%industry understand and operate networks, by changing the role of
%intermediate relays in the process of transmission of information.
%Relays are no longer limited to storing and forwarding data, but also
%take part in the coding process, through a process called recoding,
%where the relay generates new linear combinations of incoming coded
%packets without previously decoding the data. Network coding allows
%the increase of throughput, reliability, security and delay
%performance of the networks. 
%
%Compared to other broadly used coding schemes, e.g., Gallager's LDPC
%codes~\cite{gallager1962low} or Reed-Solomon
%codes~\cite{reed1960polynomial}, network coding is a technology that
%has been implemented in real systems since the early years of its
%conception. In 2006, Katabi et al.~\cite{katabi2006practical} published
%the results of the performance of their protocol COPE when implemented
%in a real wireless system, which relied on minimalistic coding to
%provide interesting gains. In 2008, Pedersen et
%al.~\cite{pedersen2008implementation} used commercially available
%Symbian OS enabled mobile phones to implement network coding in a
%\ac{D2D} cooperation-based application. In 2011, researchers at Aalborg
%University developed Kodo~\cite{kodo2011pedersen}, a C++11 network
%coding library with open source code for researchers intended to make
%network coding implementations easily available for both, the research
%community and commercial entities.
%Another case is CATWOMAN~\cite{hundeboll2012catwoman}, a protocol
%implemented on top of the BATMAN protocol~\cite{johnson2008simple} for
%IEEE 802.11 multi-hop meshed networks using some of the intuition from
%COPE, but deploying it in real systems. It is currently available as
%open source code in the Linux kernel. Many other implementations have
%been tested on real world systems~
%\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.
%
%Despite this successful deployment in real systems, many of these
%protocols and contributions have been implemented in separate testbeds
%and the experiences are hard to reproduce.  Given the need for
%reproducibility of past and future results among researchers and the
%need of low-cost and easy-to-deploy testbeds, this paper contribution
%is provide detailed instructions on how to set up a low-cost,
%heterogeneous and scalable testbed consisting of Raspberry Pis
%(Raspis), and by providing detailed measurements of goodput and energy
%consumption of such testbed when performing network coding operations
%with different codecs (such as full \ac{RLNC} \cite{ho2006random},
%multicore-enabled \ac{RLNC}, sparse \ac{RLNC} and tunable sparse \ac{RLNC}).
%The heterogeneity of the testbed comes from the fact that old, single core,
%Raspberry Pi 1 model B units are integrated seamlessly in the described
%testbed with newer, multicore versions, like the Raspberry Pi 2 model B.
%Our testbed allows us to obtain a comprehensive set of measurements
%to characterize the goodput and energetic footprint of these devices
%showing that processing speeds of 800 Mbps and energy per bit values
%of 1 nJ or even one order of magnitude less are possible.
%
%Our work is organized as follows. Section~\ref{sec:testbed} introduces the
%testbed system, its setup, files configuration and connectivity.
%Section~\ref{sec:cross_compilation} elaborates on the compilation of
%the Kodo library for the \ac{Raspi}. Section~\ref{sec:schemes} defines the
%coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
%we describe the considered metrics and methodology for performance comparison
%of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
%show the measurements in the \ac{Raspi} models of the mentioned metrics. Final
%conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
%
%%The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications should be cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the main conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.

>>>>

The upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on global scale. This will break new ground for new markets such as driverless cars, manufacturing, humanoid robots, smart grids, and many more. With 5G the number of wireless devices will increase by two fold up to 500 billion devices. It is generally accepted, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connect due to the expected energy consumption. Therefore the communication architecture in future communication scenario dominated by mesh technologies.

Mesh technology is known for ages for sensor networks, ad hoc networks, or mobile cloud scenarios, but the technical requirements on 5G mesh communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability, and heterogeneity of devices. In state of the art systems those requirements are traded off with each other, but in 5G context we cannot do this anymore.

In our previous works we have shown that random linear network coding [MURIEL] is able to satisfy the aforementioned technical requirements. We have actually shown how to increase the throughput [PEYMAN], reduce the delay [WEB1], or support heterogeneity for coding enabled communication nodes [FULCRUM]. But most of the work was focusing on small mesh networks with a handful of communication nodes. Introduced by Ahlswede et al. [1], network coding constitutes a paradigm shift in the way how researchers and industry understand and operate networks, by changing the role of intermediate relays in the process of transmission of information. Relays are no longer limited to storing and forwarding data, but also take part in the coding process, through a process called recoding, where the relay generates new linear combinations of incoming coded packets without previously decoding the data. Network coding allows the increase of throughput, reliability, security and delay performance of the networks.

Based on the need for large scale mesh communication testbeds, in the paper we show a step by step tutorial how to build up a low cost testbed with several communication nodes. The emergence of powerful and inexpensive single-board computers running full versions of a standard Operating System (OS) from many possible in the last years opens exciting possibilities for researchers. This not only allows to seamlessly develop implementations that are compatible with higher end devices and allows the reuse stable software from the Open Source community. It also enables the deployment and testing of large scale (hundreds of devices or more) distributed systems at a fraction of the cost necessary in previous years. This paper focuses on providing the critical steps and mechanisms followed to deploy, setup, and maintain a large-scale testbed based on Raspberry Pi devices of different models as well as including random linear network coding into the testbed.

%The emergence of powerful and inexpensive single-board computers
%running full versions of a standard \ac{OS} from many possible in the last
%years opens exciting possibilities for researchers. This not only
%allows to seamlessly develop implementations that are compatible with
%higher end devices and allows the reuse stable software from the Open
%Source community. It also enables the deployment and testing of large
%scale (hundreds of devices or more) distributed systems at a fraction
%of the cost necessary in previous years. This paper focuses on
%providing the critical steps and mechanisms followed to deploy, setup,
%and maintain a large-scale testbed based on Raspberry Pi devices of
%different models. 
%
%% INTRODUCTION TO NETWORK CODING
%Introduced by Ahlswede et al.~\cite{ahlswede2000network}, network
%coding constitutes a paradigm shift in the way how researchers and
%industry understand and operate networks, by changing the role of
%intermediate relays in the process of transmission of information.
%Relays are no longer limited to storing and forwarding data, but also
%take part in the coding process, through a process called recoding,
%where the relay generates new linear combinations of incoming coded
%packets without previously decoding the data. Network coding allows
%the increase of throughput, reliability, security and delay
%performance of the networks.

%Compared to other broadly used coding schemes, e.g., Gallager's LDPC
%codes~\cite{gallager1962low} or Reed-Solomon
%codes~\cite{reed1960polynomial}, network coding is a technology that
%has been implemented in real systems since the early years of its
%conception. In 2006, Katabi et al.~\cite{katabi2006practical} published
%the results of the performance of their protocol COPE when implemented
%in a real wireless system, which relied on minimalistic coding to
%provide interesting gains. In 2008, Pedersen et
%al.~\cite{pedersen2008implementation} used commercially available
%Symbian OS enabled mobile phones to implement network coding in a
%\ac{D2D} cooperation-based application. In 2011, researchers at Aalborg
%University developed Kodo~\cite{kodo2011pedersen}, a C++11 network
%coding library with open source code for researchers intended to make
%network coding implementations easily available for both, the research
%community and commercial entities.
%Another case is CATWOMAN~\cite{hundeboll2012catwoman}, a protocol
%implemented on top of the BATMAN protocol~\cite{johnson2008simple} for
%IEEE 802.11 multi-hop meshed networks using some of the intuition from
%COPE, but deploying it in real systems. It is currently available as
%open source code in the Linux kernel. Many other implementations have
%been tested on real world systems~
%\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.
%
%Despite this successful deployment in real systems, many of these
%protocols and contributions have been implemented in separate testbeds
%and the experiences are hard to reproduce.  Given the need for
%reproducibility of past and future results among researchers and the
%need of low-cost and easy-to-deploy testbeds, this paper contribution
%is provide detailed instructions on how to set up a low-cost,
%heterogeneous and scalable testbed consisting of Raspberry Pis
%(Raspis), and by providing detailed measurements of goodput and energy
%consumption of such testbed when performing network coding operations
%with different codecs (such as full \ac{RLNC} \cite{ho2006random},
%multicore-enabled \ac{RLNC}, sparse \ac{RLNC} and tunable sparse \ac{RLNC}).
%The heterogeneity of the testbed comes from the fact that old, single core,
%Raspberry Pi 1 model B units are integrated seamlessly in the described
%testbed with newer, multicore versions, like the Raspberry Pi 2 model B.
%Our testbed allows us to obtain a comprehensive set of measurements
%to characterize the goodput and energetic footprint of these devices
%showing that processing speeds of 800 Mbps and energy per bit values
%of 1 nJ or even one order of magnitude less are possible.
%
%Our work is organized as follows. Section~\ref{sec:testbed} introduces the
%testbed system, its setup, files configuration and connectivity.
%Section~\ref{sec:cross_compilation} elaborates on the compilation of
%the Kodo library for the \ac{Raspi}. Section~\ref{sec:schemes} defines the
%coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
%we describe the considered metrics and methodology for performance comparison
%of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
%show the measurements in the \ac{Raspi} models of the mentioned metrics. Final
%conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
%
%%The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications should be cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the main conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.

### 2016-06-29 testbed.tex
added franks intro

The \ac{Raspi}s in our testbed is not connected to monitors, so we
have decided to go with Raspbian lite.
%Our testbed does not have monitors so we will install Raspbian lite.
A desktop environment can be installed using the package manager
(apt-get) in case a desktop environment will be desired at a later
time.

@@

The latest Raspbian lite image can be downloaded at~\cite{raspbian}.
%To download the latest Raspbian lite, we go to the url
%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
At the time of this writing, the latest bundle was
2016-05-27-raspbian-jessie-lite.zip.

we want to proceed downloading the image in a shell instead of the browser.
to do this, first open a command shell on a linux machine. this guide is
assuming a debian-based distribution, but there should only be minor differences
if you have another package manager. i.e. the package managers and name of software packages.

After the terminal has been opened, we start by declaring a few variables to reduce
repeated typing. The first variables we declare is the Raspbian image name.
Notice that the extension has been omitted. This is because the image has been
packed into a zip file. The other variable we declare is a working directory.
This is where we will download the image to and work on it. In other words, it
will be stored in \texttt{/home/<username>/Raspbian}

@@

% Define variable]
\begin{lstlisting}[]
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

A final thing to notice is the way \$ and \# are used in the code blocks. \$
means that the command can be run with normal user permissions where \# indicates that the
command needs root permissions.
Running a command with root permissions can on most systems be done by putting
the command "sudo" in front of each line. This is illustrated in the code below
with the command "whoami" that prints the username of the caller. Lines without
leading \$ or \# is the terminal output.

@@

Next, we make the working directory and change directory (cd) into the working
directory. We can then use wget to download Raspbian lite. Unpack the image
when the download is complete.

% Download and unpack image
\begin{lstlisting}[]
$ mkdir -p ${WORKDIR}
$ cd ${WORKDIR}
$ wget http://kom.aau.dk/project/TuneSCode/raspi/${IMAGE}.zip
$ unzip ${IMAGE}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

A more recent version of Raspbian lite may be available at~\cite{raspbian}.
%\url{http://downloads.raspberrypi.org/raspbian_lite/images}.

@@

After Raspbian lite has been unpacked, there should be an .img file in the directory.
"fdisk" can be used to display the content of the image. Pass the arguments "-u sectors" to
display the sizes in sectors and "-l" to display partitions and automatically
exit fdisk again.

@@

The image is in total 2709504 sectors (1.3GiB) in size and contains two
partitions. The first partition starts at sector 8192 and
the other partition starts at sector 137216. The partitions is of type FAT32
and Linux and have sizes 63M and 1.2G. This indicates that the first partition
is most likely a boot partition and the second partition is a traditional Linux file
system. Actually, it is the Linux root filesystem (i.e. /).

>>>>

The \ac{Raspi}s in our testbed are not connected to monitors, so we
have decided to go with Raspbian lite.
%Our testbed does not have monitors so we will install Raspbian lite.
A desktop environment can be installed using the package manager
(apt-get) in case a desktop environment will be desired at a later
time.

@@

The latest Raspbian lite bundle can be downloaded from Raspbian's official
webpage~\cite{raspbian}.
%To download the latest Raspbian lite, we go to the url
%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
At the time of this writing, the latest available bundle was
\texttt{2016-05-27-raspbian-jessie-lite.zip}.

To ensure that the content of the bundle does not change, this procedure
will be based on that particular version of Raspbian lite which we have
made available at~\cite{tunescode_webpage}. All other files used in this paper
will also be available there unless specified otherwise.

%We want to proceed downloading the bundle in a shell instead of the browser.
%To do this, first open a command shell on a linux machine. This guide is
%assuming a Debian-based distribution, but there should only be minor differences
%if you have another package manager. i.e. the package managers and name of software packages.

To get started, open a Linux shell (terminal) and declare the variables below.
The \texttt{URL} and \texttt{IMAGE} specifies where the Linux bundle will be
located and \texttt{WORKDIR} specifies a working directory where we will 
store Raspbian and customize it.
Do not type in \$ and \# in the shell. Those indicates if a command can be
executed as a regular user or if root permissions are required.
Running a command as root can on most systems be done by putting
\texttt{sudo} in front of the command. This is illustrated in the
following code block 
with the command \texttt{whoami} that prints the username. Lines within a code
block without leading \$ or \# is terminal output or content of a file.

%After the terminal has been opened, we start by declaring a few variables to reduce
%repeated typing. The first variables we declare is the Raspbian image name.
%Notice that the extension has been omitted. This is because the image has been
%packed into a zip file. The other variable we declare is a working directory.
%This is where we will download the image to and work on it. In other words, it
%will be stored in \texttt{/home/<username>/Raspbian}

@@

% Define variable]
\begin{lstlisting}[]
$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%A final thing to notice is the way \$ and \# are used in the code blocks. \$
%means that the command can be run with normal user permissions where \# indicates that the
%command needs root permissions.
%Running a command with root permissions can on most systems be done by putting
%the command "sudo" in front of each line. This is illustrated in the code below
%with the command "whoami" that prints the username of the caller. Lines without
%leading \$ or \# is the terminal output.

@@

Next, make the working directory and used \texttt{cd} to change into the working
directory. Then \texttt{wget} can be used to download Raspbian lite before
unpacking the \texttt{zip} file:

% Download and unpack image
\begin{lstlisting}[]
$ mkdir -p ${WORKDIR}
$ cd ${WORKDIR}
$ wget ${URL%/*}/${IMAGE}.zip
$ unzip ${IMAGE}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%A more recent version of Raspbian lite may be available at~\cite{raspbian}.
%\url{http://downloads.raspberrypi.org/raspbian_lite/images}.

@@

After Raspbian lite has been unpacked, there should be an \texttt{img} file in the directory.
\texttt{fdisk} can be used to display the content of the image. Optionally, parse the arguments
"-u sectors" to display the sizes in sectors and "-l" to display the partitions within the
image and and automatically exit \texttt{fdisk} again:

@@

The output provides some useful knowledge about the image. First, the image is in total
2709504 sectors (1.3GiB) in size and contains two
partitions. The first partition starts at sector 8192 and
the other partition starts at sector 137216. The partitions is of type FAT32
and Linux and have sizes 63M and 1.2G. This indicates that the first partition
is most likely a boot partition and the second partition is a traditional Linux file
system. Actually, it is the Linux root filesystem (i.e. /).
